---
title: "Data Analysis with R Programming"
output: github_document
---

```{r load library, message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
```

# Week 2

## Vectors and Lists in R

### Naming and creating vectors

1. Store numeric data.
```{r}
num_vec1 = c(2.5, 48.5, 101.5)

num_vec1
```

2. Store integers (must place the letter "L" directly after each number)
```{r}
int_vec1 = c(1L, 5L, 15L)

int_vec1
```

3. Store character data.
```{r}
char_vec1 = c("Lisa", "Rose", "Jennie", "Jisoo")

char_vec1
```

4. Store logical data.
```{r}
bool_vec1 = c(TRUE, FALSE, TRUE)

bool_vec1
```

### Determining properties of vectors

Use `typeof()` function.
```{r}
typeof(num_vec1)
```

```{r}
typeof(int_vec1)
```

```{r}
typeof(char_vec1)
```

```{r}
typeof(bool_vec1)
```


We can also check if a vector is a specific type by using an **`is`** function: `is.logical()`, `is.double()`, `is.integer()`, `is.character()`. 

For the example of `is.integer()`, R will return a value of `TRUE` if the vector contains integer. 
```{r}
is.integer(int_vec1)
```

```{r}
is.integer(num_vec1)
```

### Creating lists

Lists' elements can be of any type - like dates, data frames, vectors, matrices, and more. Lists can even contain other lists.
```{r}
list_1 = list("Blackpink", 4L, 5.5, TRUE)

list_1
```

```{r}
list_2 = list(list(list(1, 3, 5)))

list_2
```

### Determining the structure of lists

Use the `str()` function.
```{r}
str(list_1)
```

```{r}
str(list_2)
```

### Naming lists

```{r}
list_3 = list('Chicago' = 1, 'New York' = 2, 'Los Angeles' = 3)

list_3

list_3$`New York`
```

## Dates and Times in R

To get the **current date**, run the `today()` function.
```{r}
today()
```

To get the **current date-time**, run the `now()` function.
```{r}
now()
```

### Converting from strings to date

```{r}
ymd(20200119)

ymd("2021-02-20")

mdy("March 21th, 2022")

dmy("22-Apr-2023")
```

### Creating date-time components

```{r}
ymd_hms("2020-01-19 20:11:59")

mdy_hm("01/20/2021 08:01")
```

### Switching between date-time & date

To convert a date-time to date, use `as_date()` function.
```{r}
now()

as_date(now())
```

### Data Frames

1. Columns should be named.
2. Data frames can include many different types of data.
3. Elements in the same column should be of the same type.

```{r}
df1 = data.frame(x = c(1, 2, 3), y = c(1.5, 5.5, 7.5))

df1
```

### Files

To create new **folder**, use `dir.create()` function. Place the name of the folder in the () of the function.
```{r, eval = FALSE}
dir.create("destination_folder")
```

To create a blank **file**, use `file.create()` function. Place the name and the type of file in the () of the function.
If the file is successfully created when the function is run, R will return a value of `TRUE` (if not, R will returm `FALSE`).
```{r, eval = FALSE}
file.create("new_text_file.txt")
file.create("new_word_file.docx")
file.create("new_csv_file.csv")
```

To **copy** a **file**, use `file.copy()` function. In the (), add the name of the file to be copied, comma, then the name of the destination folder where we want to save the copy file.
```{r, eval = FALSE}
file.copy("name_of_file.txt", "destination_folder")
```

To **delete** a file, use `unlink()` function. Enter the file name in the () of the function.
```{r, eval = FALSE}
unlink("file_name.csv")
```

## Matrices

To create matrix, use `matrix()` function. First, add a vector. Next, add at least one matrix dimension.
```{r}
# Example: Create a 2x3 (rows x columns) matrix containing the values 3-8.
matrix(c(3:8), nrow = 2)

# Example: Create a 3x2 (rows x columns) matrix containing the values 3-8.
matrix(c(3:8), ncol = 2)
```




















