---
title: "Visualization data with ggplot2"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(palmerpenguins)
```

```{r load dataset}
data(penguins)
head(penguins, 5)
glimpse(penguins)
```

Plot the relationship between body mass and flipper length in the three penguin species.

### Scatterplot

```{r}
ggplot(data = penguins) +
  geom_point(aes(x= flipper_length_mm, y = body_mass_g, shape = species, alpha = 0.8), color = 'purple')
```

```{r}
ggplot(data = penguins) +
  geom_point(aes(x= flipper_length_mm, y = body_mass_g, color = species, shape = species))
```

### Line graph

```{r}
ggplot(data = penguins) +
  geom_smooth(aes(x= flipper_length_mm, y = body_mass_g, linetype = species))
```


```{r}
ggplot(data = penguins) +
  geom_smooth(aes(x= flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point(aes(x= flipper_length_mm, y = body_mass_g, color = species))
```

### Jittering

```{r}
ggplot(data = penguins) +
  geom_jitter(aes(x= flipper_length_mm, y = body_mass_g, color = species, shape = species))
```

### Bar chart

```{r}
ggplot(data = diamonds) +
  geom_bar(aes(x = cut, fill = clarity))
```

### Facet_ functions

#### `facet_wrap()` - Separate by categories within one variable
```{r}
ggplot(data = penguins) + 
  geom_point(aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  facet_wrap(~species)
```

```{r}
ggplot(data = diamonds) +
  geom_bar(aes(x = color, fill = cut)) +
  facet_wrap(~cut)
```

#### `facet_grid()` - Separate by categories of two variables (can also do one variable)

```{r}
ggplot(data = penguins) +
  geom_point(aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  facet_grid(~sex)
```

```{r}
ggplot(data = penguins) +
  geom_point(aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  facet_grid(~species)
```

# Hands-On Activity 1

## Background for this activity

In this activity, you'll review a scenario, and use ggplot2 to quickly create data visualizations that allow you to explore your data and gain new insights. You will learn more about basic ggplot2 syntax and data visualization in R. 

Throughout this activity, you will also have the opportunity to practice writing your own code by making changes to the code chunks yourself. If you encounter an error or get stuck, you can always check the Lesson2_GGPlot_Solutions .rmd file in the Solutions folder under Week 4 for the complete, correct code. 

## The Scenario

In this scenario, you are a junior data analyst working for a hotel booking company. You have cleaned and manipulated your data, and gotten some initial insights you would like to share. Now, you are going to create some simple data visualizations with the `ggplot2` package. You will use basic `ggplot2` syntax and troubleshoot some common errors you might encounter.    

## Step 1: Import your data

```{r load data, message = FALSE}
hotel_bookings <- read_csv("./Data/hotel_bookings.csv")
```

## Step 2: Look at a sample of your data

Use the `head()` function to preview your data: 
```{r examining your data}
head(hotel_bookings)
```

You can also use `colnames()` to get the names of all the columns in your data set. Run the code chunk below to find out the column names in this data set:
```{r look at column names}
colnames(hotel_bookings)
```

## Step 3: Load the 'ggplot2' package

```{r loading and installing ggplot2, echo=FALSE, message=FALSE}
library(ggplot2) # Can load the `tidyverse` package directly.
```

## Step 4: Begin creating a plot

A stakeholder tells you, "*I want to target people who book early, and I have a hypothesis that people with children have to book in advance.*"

When you start to explore the data, it doesn't show what you would expect. That is why you decide to create a visualization to see how true that statement is-- or isn't.

You can use `ggplot2` to do this. Try running the code below:
```{r create a plot}
ggplot(data = hotel_bookings) +
  geom_point(mapping = aes(x = lead_time, y = children))
```

The geom_point() function uses points to create a scatterplot. Scatterplots are useful for showing the relationship between two numeric variables. In this case, the code maps the variable 'lead_time' to the x-axis and the variable 'children' to the y-axis. 

On the x-axis, the plot shows how far in advance a booking is made, with the bookings furthest to the right happening the most in advance. On the y-axis it shows how many children there are in a party.

The plot reveals that your stakeholder's hypothesis is incorrect. You report back to your stakeholder that many of the advanced bookings are being made by people with 0 children. 

## Step 5: Try it on your own

Next, your stakeholder says that she wants to increase weekend bookings, an important source of revenue for the hotel. Your stakeholder wants to know **what group of guests book the most weekend nights** in order to target that group in a new marketing campaign. She suggests that *guests without children book the most weekend nights*. Is this true? 

Try mapping 'stays_in_weekend_nights' on the x-axis and 'children' on the y-axis by filling out the remainder of the code below.
```{r}
ggplot(data = hotel_bookings) +
 geom_point(mapping = aes(x = stays_in_weekend_nights, y = children))
```

If you correctly enter this code, you should have a scatterplot with 'stays_in_weekend_nights' on the x-axis and 'children' on the y-axis. 

What did you discover? Is your stakeholder correct?

What other types of plots could you use to show this relationship?

Remember, if you're having trouble filling out a code block, check the solutions document for this activity. 

## Activity Wrap Up
The `ggplot2` package allows you to quickly create data visualizations that can answer questions and give you insights about your data. Now that you are a little more familiar with the basic `ggplot2` syntax, you can practice these skills by modifying the code chunks in the rmd file, or use this code as a starting point in your own project console. With `ggplot2`, you will be able to create and share data visualizations without leaving your `R` console. You will learn more about `ggplot2` throughout this course and eventually create even more complex and beautiful visualizations!


# Hands-On Activity 2

## Background for this activity
In this activity, you'll review a scenario, and continue to apply your knowledge of data visualization with ggplot2. You will learn more about the aesthetic features of visualizations and how to customize them by specific criteria. 

Throughout this activity, you will also have the opportunity to practice writing your own code by making changes to the code chunks yourself. If you encounter an error or get stuck, you can always check the Lesson3_Aesthetics_Solutions .rmd file in the Solutions folder under Week 4 for the complete, correct code. 

## The Scenario
In this example, you are a junior data analyst working for the same hotel booking company from earlier. Last time, you created some simple visualizations with `ggplot2` to give your stakeholders quick insights into your data. Now, you are are interested in creating visualizations that highlight different aspects of the data to present to your stakeholder. You are going to expand on what you have already learned about `ggplot2` and create new kinds of visualizations like bar charts.

## Making a Bar Chart

Your stakeholder is interested in developing promotions based on different booking distributions, but first they need to know **how many of the transactions are occurring for each different distribution type**.

```{r example chart 2}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel))
```

## Practice quiz
Use the bar chart you created to answer this question: what distribution type has the most number of bookings? Note your answer for the practice quiz question in Coursera afterwards.

**A: TA/TO**
B: Direct
C: GDS
D: Corporate

## Diving deeper into bar charts

After exploring your bar chart, your stakeholder has more questions. Now they want to know **if the number of bookings for each distribution type is different depending on whether or not there was a deposit or what market segment they represent**. 

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel, fill = deposit_type))
```

Now try adding 'fill = market_segment' to this code chunk instead of 'fill = deposit_type':

```{r pressure}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel, fill = market_segment))
```

## Step 6: Facets galore

After reviewing the new charts, your stakeholder asks you to create separate charts for each deposit type and market segment to help them understand the differences more clearly.

You know that the `facet_` function can do this very quickly.

Add 'deposit_type' after the '~' symbol in the code chunk below to create a different chart for each deposit type:
```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel)) +
  facet_wrap(~deposit_type)
```

This code chunk creates three bar charts for 'no_deposit', non_refund', and 'refundable' deposit types. You notice that it's hard to read the x-axis labels here, so you add one piece of code at the end that rotates the text to 45 degrees to make it easier to read. 

Try it out below:
```{r plot with rotated labels}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel)) +
  facet_wrap(~deposit_type) +
  theme(axis.text.x = element_text(angle = 45))
```
This code chunk creates a similar bar chart to the previous chunk, but now the labels on the x axis with the different distribution channels are clearer.

You can use the same syntax to create a different chart for each market segment:
```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel)) +
  facet_wrap(~market_segment) +
  theme(axis.text.x = element_text(angle = 45))
```

The `facet_grid` function does something similar. The main difference is that **`facet_grid` will include plots even if they are empty**. Run the code chunk below to check it out:

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel)) +
  facet_grid(~deposit_type) +
  theme(axis.text.x = element_text(angle = 45))
```
Now you should have three bar charts-- but notice that the 'Refundable' chart has much less data plotted than the other two. 

Now, you could put all of this in one chart and explore the differences by deposit type and market segment.

Run the code chunk below to find out; notice how the ~ character is being used before the variables that the chart is being split by: 
```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel)) +
  facet_wrap(~deposit_type~market_segment) +
  theme(axis.text.x = element_text(angle = 45))
```

These charts are probably overwhelming and too hard to read, but it can be useful if you are exploring your data through visualizations.  

## Activity Wrap Up
The `ggplot2` package allows you to create a variety of visualizations in `R`, from simple scatter plots to complicated, multi-faceted bar charts. You can practice these skills by modifying the code chunks in the rmd file, or use this code as a starting point in your own project console. As you continue exploring aesthetic arguments in `ggplot2`, consider how you might use visualizations to gain insights and make observations about other kinds of data in the future. 


# Hands-On Activity 3

## Background for this activity
In this activity, you'll review a scenario, and practice creating a data visualization with ggplot2. You will learn how to make use of the filters and facets features of ggplot2 to create custom visualizations based on different criteria. 

Throughout this activity, you will also have the opportunity to practice writing your own code by making changes to the code chunks yourself. If you encounter an error or get stuck, you can always check the Lesson3_Filters_Solutions .rmd file in the Solutions folder under Week 4 for the complete, correct code. 

## The Scenario

As a junior data analyst for a hotel booking company, you have been asked to clean hotel booking data, create visualizations with `ggplot2` to gain insight into the data, and present different facets of the data through visualization. Now, you are going to build on the work you performed previously to apply filters to your data visualizations in `ggplot2`.  

After considering all the data (in previous activities), your stakeholder decides to send the promotion to families that make online bookings for city hotels. The online segment is the fastest growing segment, and families tend to spend more at city hotels than other types of guests. 

Your stakeholder asks if you can **create a plot that shows the relationship between lead time and guests traveling with children for online bookings at city hotels**. This will give her a better idea of the specific timing for the promotion. 

You think about it, and realize you have all the tools you need to fulfill the request. You break it down into the following two steps: 1) filtering your data; 2) plotting your filtered data. 

```{r filtering a dataset to just city hotels that are online TA}
onlineta_city_hotels <- filter(hotel_bookings, 
                           (hotel == "City Hotel" & 
                             hotel_bookings$market_segment == "Online TA"))
```

Note that you can use the '&' character to demonstrate that you want two different conditions to be true. Also, you can use the '$' character to specify which column in the data frame 'hotel_bookings' you are referencing (for example, 'market_segment'). 

You can use the`View`() function to check out your new data frame:

```{r View}
View(onlineta_city_hotels)
```

There is also another way to do this. You can use the pipe operator (%>%) to do this in steps! 

You name this data frame `onlineta_city_hotels_v2`:

```{r filtering a dataset with the pipe}
onlineta_city_hotels_v2 <- hotel_bookings %>%
  filter(hotel == "City Hotel") %>%
  filter(market_segment == "Online TA")
```

Notice how in the code chunk above, the %>% symbol is used to note the logical steps of this code. First, it starts with the name of the data frame, `onlineta_city_hotels_v2`, AND THEN it tells `R` to start with the original data frame `hotel_bookings`. Then it tells it to filter on the 'hotel' column; finally, it tells it to filter on the 'market_segment' column.

This code chunk generates the same data frame by using the `View()` function:

```{r view second dataframe}
View(onlineta_city_hotels_v2)
```

You can use either of the data frames you created above for your new plots because they are the same. 

Using the code for your previous scatterplot, replace `variable_name` in the code chunk below with either `onlineta_city_hotels` or `onlineta_city_hotels_v2` to plot the data your stakeholder requested:

```{r creating a plot part two}
ggplot(data = onlineta_city_hotels) +
  geom_point(mapping = aes(x = lead_time, y = children))
```

Based on your previous filter, this scatterplot shows data for online bookings for city hotels. The plot reveals that bookings with children tend to have a shorter lead time, and bookings with 3 children have a significantly shorter lead time (<200 days). So, promotions targeting families can be made closer to the valid booking dates.  

## Activity Wrap Up

Filters allow you to create different views of your data and allow you to investigate more specific relationships within your data. You can practice these skills by modifying the code chunks in the rmd file, or use this code as a starting point in your own project console. Now that your stakeholder has had a chance to review these plots, they are interested in adding annotations they can use to explain the data in a presentation. Luckily, `ggplot2` has a function that will allow you to do just that. You will learn more about `ggplot2` in the next activity! 

# Extra Tips

* To highlight underperforming products, use an aesthetics function: `col = ifelse(x < 2, 'blue', 'yellow')`

* To create a trend line in a scatterplot, add another geom function: `geom_smooth()`.

